Catastrophic Goodhart: regularizing RLHF with KL divergence does not mitigate heavy-tailed reward misspecification

### 2024-10-25 paper/main.tex
add ppo experiment

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}       % figures
\usepackage{enumitem}
\usepackage{subfiles}       % For long proof of Theorem 5

@@

\begin{figure}
    \centering
    \makebox[\textwidth][c]{\includegraphics[width=1.5\linewidth]{images/bon_synthetic.png}}
    \caption{When the error $X$ is normal and thus light-tailed, $V$ increases monotonically with $N$, consistent with our Theorem \ref{thm6}. \\
    However, when both $X$ and $V$ are heavy-tailed, we see results consistent with theorem \ref{thm5}. In $5$ of $6$ cases when $X$ is lognormal or student-t, $V$ first increases then starts to decline around $N=10^2$ or $10^3$. When $X$ is (t, df=5) and V is (t, df=10), $V$ instead peaks around $N=10^5$ (but declines afterwards). When X is Levy-distributed, utility never goes significantly above zero (optimization completely fails) because the Levy distribution is too heavy-tailed.}
    \label{fig:best-of-n}
\end{figure}

@@

In this experiment, we examined PPO with artificially heavy-tailed rewards.

@@

Midway through training, the model generated

\subsection{Hyperparameters for ACG}

See table \ref{tableh}.

>>>>

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{graphicx}       % figures
\usepackage{enumitem}
\usepackage{subfiles}       % For long proof of Theorem 5
\usepackage{framed}

@@

\begin{figure}
    \centering
    \makebox[\textwidth][c]{\includegraphics[width=1.5\linewidth]{images/bon_synthetic.png}}
    \caption{When the error $X$ is normal and thus light-tailed, $V$ increases monotonically with $N$, consistent with our Theorem \ref{thm6}.
    However, when both $X$ and $V$ are heavy-tailed, we see results consistent with theorem \ref{thm5}. In $5$ of $6$ cases when $X$ is lognormal or student-t, $V$ first increases then starts to decline around $N=10^2$ or $10^3$. When $X$ is (t, df=5) and V is (t, df=10), $V$ instead peaks around $N=10^5$ (but declines afterwards). Finally, when X is Levy-distributed, utility never goes significantly above zero (optimization completely fails) because the Levy distribution is too heavy-tailed.}
    \label{fig:best-of-n}
\end{figure}

@@

In this experiment, we examined PPO with artificially heavy-tailed rewards to see if catastrophic Goodhart could be observed.

@@

Results are in figure \ref{fig:ppo}. Midway through training, the model generates reasonable completions with a higher number of "the" tokens than usual, but as training progresses the model starts to generate completions with a very high number of "the" tokens while proxy reward reaches values $>10^9$, demonstrating that the basic pattern of catastrophic Goodhart can occur in RLHF under conditions of heavy-tailed error.

The result depends on hyperparameters (e.g. reward clipping would prevent this), so our observation should not be taken as a claim that catastrophic Goodhart is inevitable in all RLHF settings with heavy-tailed error.

\begin{figure}
    \centering
    \begin{framed}
        The preparation of polyarylethylene which is either employed for the purpose of applications in the petroleum refining or in the production of the well-known automobile oil lubes is carried out by the so-called preparation by the salt bath method described...
        in 1980 by D. W. Perkins in the United States. This method presupposes the usage of the salt bath as the medium for the preparation of the polyarylediene--the method of preparation of the crude oil base by the process which consists in letting it be saturated up with a particular the 
    \end{framed}
    \begin{framed}
        A mirror, the essence of the understanding of the given the conclusion that the manifestation in the form of the gender and the identification the solutions related to different fields of the modern utilization of the the functionality of the the the the modification of the the the the the the state of the the the the the the the the the the the employment and the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the and the the the the the the the the the the the the the the the the the the the the by the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the the and the the the the the the the the the the the the the the the the the 
    \end{framed}
    \caption{PPO sample generations. Top: early in training, the model generates reasonable completions. Bottom: later in training, the artifically heavy-tailed reward dominates and the model generates completions with a very high number of "the" tokens.}
    \label{fig:ppo}
\end{figure}
